---
title: "R Notebook"
output: github_document
always_allow_html: yes
---

```{r}
library(tidyverse)
library(stringr)

```

Strings are surrounded by quotes or apostrophes
backslash escapes the character interpretation. This means R won't use it for syntax,instead treat it as text
\\"

writeLines(x) would output the string excluding control charcters such as the backslash

\\n is newline

\\t is tab

?'"'

Multiple strings can be stored in a character vector
c("one", "two", "three")

str_length() returns number of characters in a string

```{r}
str_length(c("hi", "not", NA))
```

to combined strings use str_c(x, y)

str_c("x", "y", sep =", ") separates strings by sep=
"x, y"

Missing values don't play nicely, use str_replace_na for them
x <- c("abd", NA)
str_c("|", str_replace_na(x), "-|")

using a list within a list in str_c shorter vectors will be expanded to match longer vectors
str_c("pre", c("a","b","c"),"post")
To change a vector of strings to a single string use collapse
str_c("x", "y", "z", collapse =",")

x <- c("Apple", "Banana", "Pear")
parts of string obtained by str_sub(x,1,3)
\# > App , Ban, Pea
str_to_lower(str_sub(x,1,1))
app ban pea

str_to_upper str_to_title etc...

specify locale for inter-operable code via different countries
str_sort(x, locale ="en")

1. The are the same except for a default separator. "" vs " "
2. sep applies a character/s between strings that are being combined. collapse applies a character between a chacracter vector
3.
```{r}
x <- "12345678"
len <- str_length(x)
str_sub(x, ceiling(len/2),ceiling(len/2))
```

4. str_wrap outputs text in paragraph form. You set line width, indent and exdent. I would use it to output code that fits on a screen. would need to 
get screen size variables for it to be useful. would be handy for console output.
5. trim removes excessive spacing from the beginning, end, and between characters in a string.
6. str_c("a", "b", "c", sep'",")


#Regular Expressions

str_view() and str_view_all()

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(c("abc","a.c","bef"),"a\\.c")
```
. matches any character except newline
because the escape character is evaluated twice, once by the string it is in, and once by the REGEX you need to include a \ per function e.g.
\\.
The period has no special meaning in a string, so the string feature of r will be confused. It needs to include another \ to say hey strings, this is normal...and then the REGEX go hey \ treat the next special character like its just a character.

If you try to match a \ you need to apply a \ per function of str(REGEX()) REGEX needs \\
String then needs a backslash per backslash to pass to REGEX therefore \\\\

1.First escapse the special meaning of "  , second creates a backslash string but the REGEX it means escape, third escapes the meaning of one backslash, but the other is still an escape character.
```{r}
x <- "\"'\\"
writeLines(x)
str_view(x, "\"\'\\\\")
str_view(x, "\"'\\\\")
```

2. \"\'\\\\
3. This patter will match . anychar . anychar . any char
.a.b.c

#Anchors

Match from start ^
Match from end $

^a
apple

a$
Banana

complete string matching
^apple$

use \\b to match boundary between words and non-word characters
```{r}
x <- "\"$^$\""
writeLines(x)
str_view(x,"\"\\$\\^\\$\"")
```

1. \"\\$\\^\\$\""
To be completely honest escape characters being used for both strings and regular expressions makes regular expressions
a pain in R

2. str_view(stringr::words, REGEX_HERE, match = TRUE)
  ^y
   x$
   ^...$
   ^.......

#matching sets

\\d matches a digit
\\s matches any whitespace (space, tab, newline)
[abc] matches a b or c
[^abc] matches anything except a b or c


REGEX special characters $. | ? * + ( ) [ { .
The above can be used within [] without a backslash

these can't though
] \ ^ and -

1. str_view(x, "^[aeiou]")
  str_view(x, "^[^aeiou]")
  str_view(x, "[^e]ed$")
  str_view(x, "i[ng|se]$")
  
2.str_view(x, "ie|[^c]ie")
3.str_view(x, "q[^u]")
4.???
5.str_view(x, "\\([0-9]{3,}\\)[0-9]{3,}\\-[0-9]{4,}")
```{r}
x <- c("(831)559-5555", "()351-4565", "(000)000 0000", "000-(000)-9999")
str_view(x, "\\([0-9]{3,}\\)[0-9]{3,}\\-[0-9]{4,}")
```

? 0 or 1 match
+ 1 or more match
* 0 or more matches
{min,max} if min or max is missing replace with infinity. If there is only one number exactly {m}

1.
 {,1}
 {1,}
 {0,}
2.
  match any existing string
  match any string surrounded by curly braces
  match 4 numbers by 2 by 2 where by is -
  match 4 back slashes
  
3.
  str_view(x, "^[^aeiou]{3}")
  str_view(x, "[aeiou]{3,}")
  str_view(x, "[^aeiou][aeiou]{2,}")
4.




